\documentclass[12pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage[no-math]{fontspec}
\setmainfont[
    BoldFont = Vollkorn Bold,
    ItalicFont = Vollkorn Italic,
    BoldItalicFont={Vollkorn Bold Italic},
    RawFeature=+lnum,
]{Vollkorn}

\setsansfont[
    BoldFont = Lato Bold,
    FontFace={l}{n}{*-Light},
    FontFace={l}{it}{*-Light Italic},
]{Lato}

\usepackage{graphicx}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{unicode-math}
\mathitalicsmode=1

\setmathfont[
mathit = sym,
mathup = sym,
mathbf = sym,
math-style = TeX, 
bold-style = TeX
]{Wholegrain Math}
\everydisplay{\Umathoperatorsize\displaystyle=4ex}
\AtBeginDocument{\renewcommand\setminus{\smallsetminus}}


\newcommand\extrafootertext[1]{%
    \bgroup
    \renewcommand\thefootnote{\fnsymbol{footnote}}%
    \renewcommand\thempfootnote{\fnsymbol{mpfootnote}}%
    \footnotetext[0]{#1}%
    \egroup
}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newenvironment{solution}
  {\textbf{Solution.} }
  {}

\usepackage{listings}
\lstset{basicstyle=\ttfamily,breaklines=true}

\usepackage{enumitem}
\setlist[itemize]{itemsep=-.15em, topsep=-.5em}
\setlist[enumerate]{itemsep=-.15em, topsep=-.5em}

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}

\setlength{\parskip}{1em}
\setlength{\parindent}{0em}

\renewcommand{\d}{\mathrm{d}}
\renewcommand{\vec}{\mathbf}

\usepackage{dsfont}
\newcommand{\bOne}{\mathds{1}}

\usepackage{float}
\usepackage{tikz}

\usepackage{booktabs}

\begin{document}

    \textbf{\Large\sffamily{Quiz 2 (bonus)\hfill Numerical Analysis Spring 2023}}
    
    \vspace{-1.8em}
    \hrulefill
\vspace{.25em} 

Answer the following questions. 
You can discuss with others, but you must write up your solutions alone.
Earning full points on this assignment will result in increasing your quiz 2 score by 7 points (with a maximum of 100\%).

\vspace{2.5em}
\begin{problem}
    Let $\mathbb{F}$ denote some discrete set of numbers, and suppose that for some $\epsilon>0$ the function $\operatorname{rd}:\mathbb{R} \to \mathbb{F}$ satisfies
    \[
        |x - \operatorname{rd}(x)|< \epsilon |x|
        ,\qquad \forall x\in\mathbb{R}.
    \]

    Find the largest value of $\epsilon$ for which we can guarantee
    $\operatorname{rd}(10^5+1) \neq \operatorname{rd}(10^5)$. 

\end{problem}

\vspace{2em}
\begin{problem}
    Consider the following problem/task: You are given a differentiable function $h:[-1,1]\to \mathbb{R}$ and must return the length 2 vector $[h'(0),\int_{-1}^{1} h(s) \d{s}]$.

            Example inputs/outputs:\vspace{-.5em}
            \begin{center}
                \begin{tabular}{ccc}
            \toprule
                    input && solution \\ \midrule
                    $h(s) = 1$ && $[0,2]$ \\
                    $h(s) = s^2+2s$ && $[2,2/3]$ \\
                    $h(s) = \sin(s)$ && $[1,0]$ \\
                \bottomrule
            \end{tabular}
            \end{center}

    Define two inputs $h$ and $\tilde{h}$ as near if $\operatorname{dist}(h,\tilde{h}):=  \int_{-1}^{1} |h(s) - \tilde{h}(s)| \d{s} $ is small.


    \begin{enumerate}[label=(\alph*)]
        \item Give a reasonable mathematical definition for the condition number of this problem at an input $h$. Your definition should be specific to this particular problem/task, not a generic answer.

        \item Decide whether this problem is well-conditioned or not. If it is, explain why. If it is not, provide an example showing that it is not (with justification). 
    \end{enumerate}

\end{problem}




\end{document}
