\documentclass[12pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{graphicx}
\usepackage{amsmath,amsthm,amssymb}

% you will probably have to remove the font-setup
\usepackage[no-math]{fontspec}
\setmainfont[
    BoldFont = Vollkorn Bold,
    ItalicFont = Vollkorn Italic,
    BoldItalicFont={Vollkorn Bold Italic},
    RawFeature=+lnum,
]{Vollkorn}

\setsansfont[
    BoldFont = Lato Bold,
    FontFace={l}{n}{*-Light},
    FontFace={l}{it}{*-Light Italic},
]{Lato}

\usepackage{unicode-math}
\mathitalicsmode=1

\setmathfont[
mathit = sym,
mathup = sym,
mathbf = sym,
math-style = TeX, 
bold-style = TeX
]{Wholegrain Math}
\everydisplay{\Umathoperatorsize\displaystyle=4ex}
\AtBeginDocument{\renewcommand\setminus{\smallsetminus}}
% font setup ends here

\newcommand\extrafootertext[1]{%
    \bgroup
    \renewcommand\thefootnote{\fnsymbol{footnote}}%
    \renewcommand\thempfootnote{\fnsymbol{mpfootnote}}%
    \footnotetext[0]{#1}%
    \egroup
}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newenvironment{solution}
  {\textbf{Solution.} }
  {}

\usepackage{listings}
\lstset{basicstyle=\ttfamily,breaklines=true}

\usepackage{enumitem}
\setlist[itemize]{itemsep=-.15em, topsep=-.5em}
\setlist[enumerate]{itemsep=-.15em, topsep=-.5em}
\setlist[enumerate,1]{label={(\alph*)}}

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}

\setlength{\parskip}{1em}
\setlength{\parindent}{0em}

\renewcommand{\d}{\mathrm{d}}
\renewcommand{\vec}{\mathbf}
\newcommand{\T}{\mathsf{T}}

\usepackage{dsfont}
\newcommand{\bOne}{\mathds{1}}


\begin{document}

\textbf{\Large\sffamily{Intro Survey\hfill Numerical Analysis Fall 2024}}
    
    \vspace{-1.8em}
    \hrulefill
 
\textbf{Instructions:}
    \begin{itemize}
        \item Due 09/10 at 5:00pm on \href{https://www.gradescope.com/courses/818054}{Gradescope}.
        \item You must follow the submission policy in the \href{https://courses.chen.pw/na_f2024/syllabus.html}{syllabus} 
\end{itemize}
   
\vspace{2em}

\begin{problem}
    ~
    \begin{itemize}
        \item Respond to the intro post on Ed. If you do not have access, contact me as soon as possible.
        \item What percent of the final grade is participation? What do you need to do to get full participation credit?
        \item Check if you have any conflicts which will force you to miss class. If you do, please let me know now.
            Examples of common conflicts include midterms for other classes which are scheduled during our class time, standardized tests, religious holidays, etc.
            Quizzes cannot be made up if you miss class due to a foreseeable absence and did not notify me sufficiently far in advance.
    \end{itemize}
\end{problem}

\begin{problem}
Describe in a few sentences your background with: 
    \begin{itemize}
        \item linear algebra
        \item coding
    \end{itemize}
\end{problem}

\begin{problem}
Tell me about your goals for the class.
    \begin{itemize}
        \item Are there any particular things you are hoping to learn? 
        \item What are you most excited about?
        \item Is there anything you're nervous about?
        \item Is there anything else you want me to know?
    \end{itemize}
\end{problem}


\end{document}
